Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-20 main.tex
corrections

\begin{remark}
    The ratio $\aspectratio(T)$ measures the ``shape quality'' of an $n$-dimensional simplex $T$ and is an instance of a so-called \emph{shape measure}.
    For example, the reference triangle has aspect shape measure $2$ and the reference tetrahedron has aspect shape measure $\sqrt{6}$. 
    Numerous alternative shape measures have been used throughout the literature of numerical analysis and computational geometry to quantify the quality of simplices (see~\cite[p.61, Definition 5.1]{braess2001finite}, 
    \cite[p.97, Definition 4.2.16 ]{brenner2008mathematical}, % TODO: add brackets here in the final version or figure out how to fix the display in the editor.
    \cite[Definition~11.2]{ern2021finite}). 
\end{remark}

@@

    We let $\hat u := \varphi^{\ast} u \in W^{p}\Lambda^{k}(\Delta^n)$ and define $\hat g \in L^{p}\Lambda^{k+1}(\Delta^n)$ via $\hat g := \cartan \varphi^{\ast} u$. 
    Then $\cartan \hat u = \hat g$. 
    We let $\tilde u \in W^{p}\Lambda^{k}(\widetilde \Makoto)$ be the extension of $\hat u$ onto $\widetilde \Makoto$ by reflection across the coordinate axes,
    and let $\tilde g \in L^{p}\Lambda^{k+1}(\widetilde \Makoto)$ is the extension of $\hat g$ onto $\widetilde \Makoto$ by reflection across the coordinate axes. 
    By construction, $\tilde u \in W^{p}_{0}\Lambda^{k}(\widetilde \Makoto)$ with $\tilde g = \cartan \tilde u$.
    We observe 
    \begin{align*}
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        \leq 
        2^{\frac {n-\ell} p}
        \| \hat g \|_{L^{p}(\Delta^n)}
        .
    \end{align*}
    By the analysis for the Bogovski\u{\i} potential operators, Theorem~\ref{theorem:bogovpoinc}, 
    there exists $\mathring w \in W^{p}_{0}\Lambda^{k}(\widetilde \Makoto)$
    with $\cartan \mathring w = \cartan \tilde u$ satisfying the bounds 
    \begin{align*}
        \| \mathring w \|_{L^{p}(\widetilde \Makoto)}
        &\leq 
        C_{{\PF},\Bogov,k,\widetilde \Makoto,p}
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        %\\&
        \leq 
        % n C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_2(0)) }{ \vol(\widetilde \Makoto) } 2
        % 2^n n   C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_1(0))    }{ \vol(\widetilde \Makoto) } 2
        % 2^n     C_{\Bogov}(n,k+1) \frac{ \vol_{n-1}(S_1(0)) }{ \vol(\widetilde \Makoto) } 2
        % 2^{n+1} C_{\Bogov}(n,k+1) \frac{ \vol_{n-1}(S_1(0)) }{ \vol(\widetilde \Makoto) }
        % 2^{n+1} C_{\Bogov}(n,k+1) \frac{ \vol_{n-1}(S_1(0)) }{ 2^{n-\ell}/n! }
        %
        % n! 2^{\ell+1} C_{\Bogov}(n,k+1) \vol_{n-1}(S_1(0)) 
        %
        n! \cdot 2^{\ell+1} C_{\Bogov}(n,k+1) \vol_{n-1}(S_1(0)) 
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        .
    \end{align*}
    Here,
    \begin{align*}
        C_{{\PF},\Bogov,k,\widetilde \Makoto,p} 
        &
        = 
        C_{\Bogov}(n,\mia) \vol_{n-1}(S_{1}(0)) \frac{\diam(\widetilde \Makoto)^{n}}{\vol(\widetilde \Makoto)} \diam(\widetilde \Makoto)
        \\&
        = 
        C_{\Bogov}(n,\mia) \vol_{n-1}(S_{1}(0)) \frac{2^{n}}{2^{n-\ell}/n!} 2
        = 
        C_{\Bogov}(n,\mia) \vol_{n-1}(S_{1}(0)) \cdot n! 2^{\ell+1} 
        .
    \end{align*}
    We let $\tilde w \in W^{p}\Lambda^{k}(\widetilde \Makoto)$ be defined by averaging the $2^{n - \ell}$ reflections of $\mathring w$ along the coordinate axes $\ell+1$ through $n$. Thus,
    \begin{align*}
        \| \tilde w \|_{L^{p}(\Delta^{n})}
        =
        2^{-\frac{n-\ell}{p}}
        \| \tilde w \|_{L^{p}(\widetilde \Makoto)}
        ,
        \qquad 
        \| \tilde w \|_{L^{p}(\widetilde \Makoto)}
        \leq 
        \| \mathring w \|_{L^{p}(\widetilde \Makoto)}
        .
    \end{align*}
    Since $\tilde g$ has got the same symmetries, $\cartan \tilde w = \tilde g$.

@@

\begin{remark}\label{remark:mixedbconsimplex:exteriorderivative:hilbert}
    Under the assumptions of Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative}, 
    the special case $p=2$ allows for the following improvement:
    there exists $w \in W^{2}\Lambda^{k}(T)$ such that 
    $\trace_{T,\Gamma} w = 0$
    and  
    \begin{align*}
        \cartan w = \cartan u,
        \quad 
        \| w \|_{L^{p}(\Omega)} 
        \leq 
        C_{\PF,\Gamma,\ell,k,2}(T)
        \| \cartan u \|_{L^{p}(\Omega)}
        ,
    \end{align*}
    where 
    \begin{align*}
        C_{\PF,\Gamma,\ell,k,p}(T)
        \leq 
        \frac{1}{\pi}
        \algebraicshapemeasure(T)^{k-1}
        \Ceins{n}(T) 
        \diam(T)
        .
    \end{align*}
    This follows from the results in \cite{licht2024geometry} as simplices are Lipschitz domains.
    We conjecture that the constant can be improved further, to be independent of the tetrahedron's eccentricity.
\end{remark}

@@

% TODO: Keep this theorem in favor of a broad audience? Or toss it?
\begin{theorem}\label{theorem:poincarefriedrichsestimate:curl}
    Let $\calT$ be a shellable $n$-dimensional triangulation, 
    and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Assume that $T_0, T_1, \dots, T_M$ be a shelling of $\calT$ and that $\bfu \in \bfW^{p}(\Omega,\curl)$ for some $1 \leq p \leq \infty$.
    Let $0 \leq \ell_{m} < n$ be such that $T_{m}$ shares $n - \ell_{m}$ faces with the previous simplices.
    \begin{itemize}
        \item 
        There exists $\bfw \in \bfW^{p}(\Omega,\curl)$ with $\curl \bfw = \curl \bfu$
        and such that the following estimates hold:
        \begin{gather*}
            \| \bfw \|_{L^{p}(T_{0})} \leq C_{{\PF},1,T_{0},p} \| \curl \bfu \|_{L^{p}(T_{0})}
        \end{gather*}
        and for each $T_{m} \in \calT$, $1 \leq m \leq M$ we have the recursive estimate 
        \begin{align*}
            \| \bfw \|_{L^{p}(T_{m})}
            &
            \leq  
            C_{\PF,T_{m},\Gamma_{m},1,p}(\calT) 
            \left( 
                \| \curl \bfu      \|_{L^{p}(T_{m})} 
                +
                \Cfivetrafo{n}{1}{\ell_{m}}{p}(\calT)
                \| \curl \bfu \|_{L^{p}(\Makoto_{m-1})}
            \right)
            \\&\qquad
            + 
            \Cfivetrafo{n}{1}{\ell_{m}}{p}(\calT)
            \| \bfw \|_{L^{p}(\Makoto_{m-1})}
            .
        \end{align*}
        \item 
        There exists $\bfw \in \bfW^{p}(\Omega,\curl)$ with $\curl \bfw = \curl \bfu$ 
        such that 
        \begin{align*}
            \| \bfw \|_{L^{p}(T_{m})}
            \leq 
            \left( 
            \prod_{i=1}^{M}
            \Cachttrafo{n}{2}{\ell_{m}}{p}(\calT)
            \Csiebtrafo{n}{1}{\ell_{m}}{p}(\calT)
            \right)
            C_{{\PF},1,T_{0},p}
            \| \curl \bfu \|_{L^{p}(\Omega)}
            .
        \end{align*}
    \end{itemize}
\end{theorem}

@@

\begin{table}[t]
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
                           & $\grad$           & $\curl$           & $\divergence$   \\ \hline
        $l=0$              & 0.08625017654     & 0.1453729386      & 0.2601720480    \\ \hline
        $l=1$              & 0.1093817645      & 0.1829680131      & 0.3493931507    \\ \hline
        $l=2$              & 0.1450219664      & 0.2428248005      & 0.1845090722    \\ \hline
        $l=3$              & 0.2057601732      & 0.1527746636      & 0.1223402289    \\ \hline
        $\Gamma=\emptyset$ & 0.2631059409      & 0.1458215887      & 0.08740665548   \\ \hline
    \end{tabular}
    \caption{
    Reference $L^2$ Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators over the reference tetrahedron,
    where boundary conditions are imposed along the faces $F_{\ell}, \dots, F_3$, or no boundary conditions are imposed (last row). 
    Computed via lowest-order finite element methods. The upper bound in Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative} is $0.45015815807$.
    }\label{table:referencepoincarefriedrichs}
\end{table}

@@

\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
    Domain & Exact PF & Approximate PF & Ratio 
    \\ 
    \hline
    $\Omega$ & $1.0$ & $1.0$ & $1.0$ 
    \\ 
    %\hline
    $\Omega$ & $1.0$ & $1.0$ & $1.0$ 
    \\ 
    %\hline
    $\Omega$ & $1.0$ & $1.0$ & $1.0$ 
    \\ 
    \hline
    \end{tabular}
    \caption{Example table with text and formulas.}
    \label{tab:example}
\end{table}

@@

The reference Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators are found via standard finite element eigenvalue computations, using the lowest-order finite element de~Rham complex over a sufficiently refined mesh (four steps of uniform refinement).
Again, we compare these reference values with estimates obtained Theorem~\ref{theorem:poincarefriedrichsestimate:exterior} and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}, and in the special case of the gradient, with Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
    Domain & Exact PF & Approximate PF & Ratio 
    \\ 
    \hline
    $\Omega$ & $1.0$ & $1.0$ & $1.0$ 
    \\ 
    %\hline
    $\Omega$ & $1.0$ & $1.0$ & $1.0$ 
    \\ 
    %\hline
    $\Omega$ & $1.0$ & $1.0$ & $1.0$ 
    \\ 
    \hline
    \end{tabular}
    \caption{Example table with text and formulas.}
    \label{tab:example}
\end{table}
The reference Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators are found via standard finite element eigenvalue computations, using the lowest-order finite element de~Rham complex and a sufficiently refined mesh (four steps of uniform refinement).
As in the two-dimensional case outlined above, we again compare these reference values with estimates obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad}, Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}, and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.

@@

This greedy backtracking algorithm does generally not guarantee that the optimal shelling is found before the last iteration. 
Our implementation starts with a fixed simplex and generates up to $10$ shellings. 
We repeat this for a possible choices of initial simplex, and pick the optimal among those.
The candidate shellings that we obtain in this manner allow for estimates that differ by a factor of about $2$.

\todo{How to select shellings}

>>>>

\begin{remark}
    The ratio $\aspectratio(T)$ measures the ``shape quality'' of an $n$-dimensional simplex $T$ and is an instance of a so-called \emph{shape measure}.
    For example, the reference triangle has aspect shape measure $2$ and the reference tetrahedron has aspect shape measure $\sqrt{6}$. 
    Numerous alternative shape measures have been used throughout the literature of numerical analysis and computational geometry to quantify the quality of simplices (see~\cite[p.61, Definition 5.1]{braess2001finite}, 
    \cite[p.97, Definition {\char`\(}4.2.16{\char`\)} ]{brenner2008mathematical}, % DONE: add brackets here in the final version or figure out how to fix the display in the editor.
    \cite[Definition~11.2]{ern2021finite}). 
\end{remark}

@@

    We let $\hat u := \varphi^{\ast} u \in W^{p}\Lambda^{k}(\Delta^n)$ and define $\hat g \in L^{p}\Lambda^{k+1}(\Delta^n)$ via $\hat g := \cartan \varphi^{\ast} u$. 
    Then $\cartan \hat u = \hat g$. 
    We let $\tilde u \in W^{p}\Lambda^{k}(\widetilde \Makoto)$ be the extension of $\hat u$ onto $\widetilde \Makoto$ by reflection across the coordinate axes,
    and let $\tilde g \in L^{p}\Lambda^{k+1}(\widetilde \Makoto)$ is the extension of $\hat g$ onto $\widetilde \Makoto$ by reflection across the coordinate axes. 
    By construction, $\tilde u \in W^{p}_{0}\Lambda^{k}(\widetilde \Makoto)$ with $\tilde g = \cartan \tilde u$.
    We observe 
    \begin{align*}
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        \leq 
        2^{\frac {n-\ell} p}
        \| \hat g \|_{L^{p}(\Delta^n)}
        .
    \end{align*}
    By the analysis for the Bogovski\u{\i} potential operators, Theorem~\ref{theorem:bogovpoinc}, 
    there exists $\mathring w \in W^{p}_{0}\Lambda^{k}(\widetilde \Makoto)$
    with $\cartan \mathring w = \cartan \tilde u$ satisfying the bounds 
    \begin{align*}
        \| \mathring w \|_{L^{p}(\widetilde \Makoto)}
        &\leq 
        C_{{\PF},\Bogov,k,\widetilde \Makoto,p}
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        %\\&
        \leq 
        % n C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_2(0)) }{ \vol(\widetilde \Makoto) } 2
        % 2^n n   C_{\Bogov}(n,k+1) \frac{ \vol_{}(B_1(0))    }{ \vol(\widetilde \Makoto) } 2
        % 2^n     C_{\Bogov}(n,k+1) \frac{ \vol_{n-1}(S_1(0)) }{ \vol(\widetilde \Makoto) } 2
        % 2^{n+1} C_{\Bogov}(n,k+1) \frac{ \vol_{n-1}(S_1(0)) }{ \vol(\widetilde \Makoto) }
        % 2^{n+1} C_{\Bogov}(n,k+1) \frac{ \vol_{n-1}(S_1(0)) }{ 2^{n-\ell}/n! }
        %
        % n! 2^{\ell+1} C_{\Bogov}(n,k+1) \vol_{n-1}(S_1(0)) 
        %
        n! \cdot 2^{\ell+1} C_{\Bogov}(n,k+1) \vol_{n-1}(S_1(0)) 
        \| \tilde g \|_{L^{p}(\widetilde \Makoto)}
        .
    \end{align*}
    Here,
    \begin{align*}
        C_{{\PF},\Bogov,k,\widetilde \Makoto,p} 
        &
        = 
        C_{\Bogov}(n,\mia+1) \vol_{n-1}(S_{1}(0)) \frac{\diam(\widetilde \Makoto)^{n}}{\vol(\widetilde \Makoto)} \diam(\widetilde \Makoto)
        \\&
        = 
        C_{\Bogov}(n,\mia+1) \vol_{n-1}(S_{1}(0)) \frac{2^{n}}{2^{n-\ell}/n!} 2
        = 
        C_{\Bogov}(n,\mia+1) \vol_{n-1}(S_{1}(0)) \cdot n! 2^{\ell+1} 
        .
    \end{align*}
    We let $\tilde w \in W^{p}\Lambda^{k}(\widetilde \Makoto)$ be defined by averaging the $2^{n - \ell}$ reflections of $\mathring w$ along the coordinate axes $\ell+1$ through $n$. Thus,
    \begin{align*}
        \| \tilde w \|_{L^{p}(\Delta^{n})}
        =
        2^{-\frac{n-\ell}{p}}
        \| \tilde w \|_{L^{p}(\widetilde \Makoto)}
        ,
        \qquad 
        \| \tilde w \|_{L^{p}(\widetilde \Makoto)}
        \leq 
        \| \mathring w \|_{L^{p}(\widetilde \Makoto)}
        .
    \end{align*}
    Since $\tilde g$ has got the same symmetries, $\cartan \tilde w = \tilde g$.

@@

\begin{remark}\label{remark:mixedbconsimplex:exteriorderivative:hilbert}
    Under the assumptions of Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative}, 
    the special case $p=2$ allows for the following improvement:
    there exists $w \in W^{2}\Lambda^{k}(T)$ such that 
    $\trace_{T,\Gamma} w = 0$
    and  
    \begin{align*}
        \cartan w = \cartan u,
        \quad 
        \| w \|_{L^{p}(\Omega)} 
        \leq 
        C_{\PF,\Gamma,\ell,k,2}(T)
        \| \cartan u \|_{L^{p}(\Omega)}
        ,
    \end{align*}
    where 
    \begin{align*}
        C_{\PF,\Gamma,\ell,k,p}(T)
        \leq 
        \frac{2}{\pi}
        \algebraicshapemeasure(T)^{k-1}
        \Ceins{n}(T) 
        \diam(T)
        .
    \end{align*}
    This follows from the results in \cite{licht2024geometry} as simplices are Lipschitz domains.
    We conjecture that the constant can be improved further, to be independent of the tetrahedron's eccentricity.
\end{remark}

@@

@@

\begin{table}[t]
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
                           & $\grad$           & $\curl$           & $\divergence$   \\ \hline
        $l=0$              & 0.08625017654     & 0.1453729386      & 0.2601720480    \\ \hline
        $l=1$              & 0.1093817645      & 0.1829680131      & 0.3493931507    \\ \hline
        $l=2$              & 0.1450219664      & 0.2428248005      & 0.1845090722    \\ \hline
        $l=3$              & 0.2057601732      & 0.1527746636      & 0.1223402289    \\ \hline
        $\Gamma=\emptyset$ & 0.2631059409      & 0.1458215887      & 0.08740665548   \\ \hline
    \end{tabular}
    \caption{
    Reference $L^2$ Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators over the reference tetrahedron,
    where boundary conditions are imposed along the faces $F_{\ell}, \dots, F_3$, or no boundary conditions are imposed (last row). 
    Computed via lowest-order finite element methods. 
    Notably, the top-left should and the bottom-right should asymptotically coincide, as should the bottom-left and the top-right.
    The upper bound in Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative} is $0.45015815807$.
    }\label{table:referencepoincarefriedrichs}
\end{table}

@@

@@

The reference Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators are found via standard finite element eigenvalue computations, using the lowest-order finite element de~Rham complex over a sufficiently refined mesh (four steps of uniform refinement).
Again, we compare these reference values with estimates obtained Theorem~\ref{theorem:poincarefriedrichsestimate:exterior} and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}, and in the special case of the gradient, with Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.

% The reference Poincar\'e--Friedrichs constants for the gradient, curl, and divergence operators are found via standard finite element eigenvalue computations, using the lowest-order finite element de~Rham complex over a sufficiently refined mesh (four steps of uniform refinement).
% As in the two-dimensional case outlined above, we again compare these reference values with estimates obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad}, Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}, and Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.

@@

This greedy backtracking algorithm does generally not guarantee that the optimal shelling is found until all shellings have been produced.
However, finding all shellings seems practically infeasible. 
Our implementation starts with a fixed simplex and generates up to $10$ shellings. 
We repeat this for a possible choices of initial simplex, and pick the optimal shelling among those found.
The candidate shellings that we obtain in this manner allow for estimates that differ by a factor of about $2$.
